---
title: "Assignment 9"
author: "Janessa McIlnay"
output: html_document
date: "April 15, 2025"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(ggplot2)
```

## Introduction

In this report, I will analyze predictors of graduate school admissions! The dataset includes:

- `admit`: Admission outcome (1 = admitted, 0 = not admitted)
- `gre`: GRE score
- `gpa`: Undergraduate GPA
- `rank`: Rank of the undergraduate institution (1 = highest)

Let's set up! 


```{r libraries, echo=TRUE, warning=FALSE}
library(tidyverse)
library(easystats)
library(caret)

grad <- read.csv('GradSchool_Admissions.csv')
```

## Analyzing the Data

Let's see how the data is structured and change it if needed!

```{r str, echo=TRUE}
str(grad)
as.logical(grad$admit)
```

Okay, let's make some models to see how the data works!

```{r modeling, echo=TRUE}
mod1 <- glm(data = grad,
            formula = admit ~ gre + gpa,
            family = 'binomial')
mod2 <- glm(data = grad,
            formula = admit ~ (gre + gpa)*rank,
            family = 'binomial')
mod3 <- glm(data = grad,
            formula = admit ~ gre*gpa*rank,
            family = 'binomial')
```

Great, now lets see how they do!

```{r testmods, echo=TRUE}
compare_performance(mod1, mod2, mod3) %>% plot()
```
Hmmm, it looks like they are all not doing great at finding correlation between variables. Let's try again.

```{r explore}

# Admission rate by rank
admission_by_rank <- grad %>%
  group_by(rank) %>%
  summarise(admit_rate = mean(admit), count = n())
print(admission_by_rank)
```
It looks like people with a rank of '1' get admitted most often. Let's make some more plots. 

# Visualizing the Data
```{r plots}

ggplot(grad, aes(x = gre, fill = as.factor(admit))) + 
  geom_histogram(position = "identity", alpha = 0.6, bins = 30) +
  labs(title = "GRE Scores and Admission Status", fill = "Admitted")
```
This graph shows that most individuals who are admitted to graduate school seem to have a gre between 500-700.

```{r plot again}
ggplot(grad, aes(x = rank, fill = as.factor(admit))) +
  geom_bar(position = "fill") +
  labs(title = "Admission Rate by Institution Rank", fill = "Admitted") +
  ylab("Proportion")
```
In this graph, it seems that higher ranking individuals are admitted more than those of a lower rank. Let's make one last graph to see how GPA affects admission.

```{r last plot}
ggplot(grad, aes(x = gpa, fill = as.factor(admit))) + 
  geom_histogram(position = "identity", alpha = 0.6, bins = 30) +
  labs(title = "GPA by Admission Status", fill = "Admitted")
```
It seems as though there is not a clear correlation between GPA and admission,  however, those with a GPA around 3.5-4.0 seem to be admitted most often.

## Model Fitting: Logistic Regression

```{r logistic-model}
# Split data
set.seed(123)
trainIndex <- createDataPartition(grad$admit, p = .8, list = FALSE)
train <- grad[trainIndex, ]
test <- grad[-trainIndex, ]

# Fit logistic regression
model <- glm(admit ~ gre + gpa + rank, data = train, family = "binomial")
summary(model)
```

This looks confusing, but it is actually very helpful for understanding the data. Let's break it down!

Gre is NOT statistically significant, as the p value is 0.16. This means it cannot be used to predict admission. Rank has a negative coefficient, meaning that having a lower rank hurts chances of admission. Additionally, the positive gpa coefficient shows that having a high gpa increases one's chances of acceptance. 

So...gpa and rank are significant admission predictors.

## New Model

```{r evaluation}
mod4 <- glm(data = grad,
            formula = admit ~ rank * gpa,
            family = 'binomial')
summary(mod4)
compare_performance(mod1, mod2, mod3, mod4) %>% plot
```

## Conclusion

After this analysis, I have determined that model 4 is the best at predicting admission. Though the r2 is still low, the graphs created show that having a higher gpa and a better rank typically lead to higher admission rates. 
